---
title: "Wrangling Data in R (Katacoda Scenario)"
output: 
  github_document: 
    fig_height: 6
    fig_width: 8
    toc: yes
---

```{r setup, include=FALSE}
library(hrbrthemes)
library(extrafont)
library(knitr)
library(rmdformats)
library(tidyverse)
library(magrittr)
library(skimr)
library(rlang)
library(gt)
library(visdat)
library(inspectdf)
library(foreign)
library(naniar)
# images folder
if (!file.exists("figs")) {
  dir.create("figs")
}
knitr::opts_chunk$set(
  echo = TRUE,
  eval = FALSE,
  cache = FALSE,
  prompt = FALSE,
  tidy = FALSE,
  fig.width = 10,
  fig.height = 7.5,
  comment = "#>",
  message = FALSE,
  warning = FALSE,
  fig.path = "figs/"
)
knitr::opts_knit$set(
  width = 78
)
base::options(
  tibble.print_max = 25,
  tibble.width = 78,
  scipen = 100000000
)
ggplot2::theme_set(hrbrthemes::theme_ipsum_tw(
  base_size = 10,
  strip_text_size = 11,
  axis_title_size = 13,
  plot_title_size = 17,
  subtitle_size = 13,
  base_family = "Ubuntu",
  strip_text_family = "TitilliumWeb-Regular",
  axis_title_family = "TitilliumWeb-Regular",
  subtitle_family = "TitilliumWeb-Regular",
  plot_title_family = "JosefinSans-Regular"
))
```

## Setup  

This scenario is built using the following commands in the Terminal pane in RStudio. 

```bash
# create a new scenario
$ katacoda scenarios:create
? Friendly url:  1-wrangle-data-in-r
? Scenario Title:  Wrangling Data in R
? Scenario Description:  Learn how to restructure and manipulate data in R
? Difficulty Level:  Intermediate
? Estimated time: (Please specify in minutes or hours e.g. 2 hours) 90 minutes
? Number of Steps (Not including intro & finish):  24
? Image:  rlang:3.4
? Layout:  Editor + Terminal
Scenario created successfully.
```

The following folder was created from these commands. 

```{r 1-wrangle-data-in-r}
fs::dir_tree("1-wrangle-data-in-r")
```

We can see all `steps`, the `intro`, and `finish` markdown files. There is also an `index.json` file for managing all the steps in the Katacoda environment.

## Markdown syntax (Katacoda)

Katacoda comes with it's own flavor of Markdown, which you can learn more about [here.](https://katacoda.com/scenario-examples/scenarios/markdown-extensions). The main thing to know is that there are `execute` and `copy` extensions, which you can use in the markdown to indicate whether Katacoda should copy or run the code inside the R REPL.

## Learner personas

The learner personas for this scenario are: 

1. Horatio is an executive at a tech company...

2. Judy is a product manager at a start-up...

3. Andrew is a graduate student in college...

***

### Intro

This scenario will introduce you to the R language, data structures, the `tidyverse`, and it's underlying philosophy. 

how to change columns from one type (or class) to another. In the wild, we rarely collect data in a format that they can be immediately analyzed or visualized in, so data wrangling is a common task.

#### Variable types

Recall that variables can be characters (or strings), logicals (or boolean), and numbers (integer and double). In this first exercise, we'll be converting variables from one form to another using [`dplyr`](https://dplyr.tidyverse.org/), a package from the [`tidyverse`](https://www.tidyverse.org/).

### step 1

The `step1` file starts the scenario off an intro to R packages (the `library()` function). The markdown for the `step1.md` file is below. 

~~~

~~~


## step 2



## step 3



## Appendix 1: Katacoda scenario tutorials

The tutorial for building the scenario is [here](https://katacoda.com/scenario-examples/scenarios/create-scenario-101).

I took notes on this tutorial and made them available in [this Google document](https://docs.google.com/document/d/e/2PACX-1vSf2w2onhH5t3IhuD4sYLoWqn46BLKMYFR7q3BHO8QTaRkVgXfhKvnl8T9uHrjmbVpTZVKCWrfxEl0R/pub).

## Appendix 2: Katacoda guidelines

There are multiple guidelines and resources for writing scenarios. O'Reilly has provided an *Authoring Guide* and *Formatting and Design Guide*.

The link for these files can be found here:

+ [Katacoda Scenario Formatting and Design Guide for Authors](https://docs.google.com/document/d/1l4lofG5kAu8JFzumZPCsJJE2muCYe6rHSHCQsMlijd8/edit)

+ [Katacoda Scenario Authoring Guide](https://docs.google.com/document/d/14rudtruZQhRxvD3zcR3g75j5nuOgKGz4CYk8hdhaV-w/edit)

## Appendix 3: Scenario Checklist

**Scenario Readiness Checklist:**

When you think your scenario is ready for publication, we recommend you run through the checklist below to ensure it is ready to go. [Reminder: We discuss these best practices in the Katacoda Formatting and Design Guide for Authors]:

Does the scenario start consistently and in a timely manner? More than 5 minutes to start would be cause to reconsider your build decisions.

- [ ] Are your intro and final pages present and are their respective goals and lessons learned in agreement?

- [ ] Have you tested your scenario lately?

- [ ] Do you agree with the leveling (beginner, intermediate, advanced) you indicated when you started the build?

- [ ] Was learning time you entered correct, or should it be adjusted?

- [ ] Are your versions of tools and other dependencies up to date?

- [ ] Have you tried every instruction?

- [ ] Have you written each step in the most concise manner possible?

- [ ] Have you run your text through a spelling/grammar checker?

- [ ] Are your credits to others given present and correct?

- [ ] Are your images legal and with credits?

- [ ] Are your hyperlinks all working?

- [ ] Do the goals and lessons learned items match the steps in the scenario?


## Scenario contents

This scenario covers...
